Beating Simple Backtracking Search in Subgraph Algorithms\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1].}

### 2018-11-29 paper/paper.tex
More little tweaks

\begin{abstract}
    The current state of the art in subgraph isomorphism solving involves using degree as a
    value-ordering heuristic to direct backtracking search. Such a search makes a heavy commitment
    to the first branching choice, which is often incorrect. To mitigate this, we introduce and
    evaluate a new approach, which we call ``solution-biased search''. By combining a
    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we design an
    algorithm which instead uses degree to direct the proportion of search effort spent in different
    subproblems. This increases performance by two orders of magnitude on satisfiable instances,
    whilst not affecting performance on unsatisfiable instances. This algorithm can also be
    parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
    instances, we get a further speedup of over thirty from thirty six cores, or ninety five from
    ten distributed-memory hosts. Finally, we show that solution-biased search is also suitable for
    optimisation problems, by using it to improve two maximum common induced subgraph algorithms.
\end{abstract}

@@

This new form of search can also be parallelised, with a \emph{much} simpler implementation than
conventional work-stealing. By running many threads with different random seeds but the same restart
schedule, and sharing nogoods only following restarts, we can achieve speedups of thirty one from a
thirty six core machine, or ninety five by using ten such machines with message passing.

\subsection{Definitions}

@@

\subsection{The Glasgow Subgraph Solver}

@@

\Cref{figure:parallel} shows how this scheme performs in practice. Sequentially, we can solve 14,357
instances within the thousand second timeout, with the last instance being solved at 939.0 seconds.
Using thirty six threads on machines with two eighteen core processors, we can solve 14,357
instances with a timeout of only 74.2 seconds, giving an aggregate speedup
\cite{DBLP:conf/cpaior/HoffmannMNPRS018} of 12.7.

@@

Because we are using nogood recording, an alternative approach is possible. Rather than using the
Luby sequence for restarts, we could restart after either a constant number of backtracks, or simply
after a certain time interval has passed---this bounds the maximum possible idle time that threads
could spend at a barrier. \Cref{figure:parallel} also shows the effects of restarting every 100ms.
Sequentially, this approach is slightly better than using the Luby sequence, being able to solve
14,370 instances, with the last at 996.4 seconds. With thirty six threads, solving this many
instances takes 31.8 seconds, giving an aggregate speedup of 31.4.

>>>>

\begin{abstract}
    The current state of the art in subgraph isomorphism solving involves using degree as a
    value-ordering heuristic to direct backtracking search. Such a search makes a heavy commitment
    to the first branching choice, which is often incorrect. To mitigate this, we introduce and
    evaluate a new approach, which we call ``solution-biased search''. By combining a
    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we design an
    algorithm which instead uses degree to direct the proportion of search effort spent in different
    subproblems. This increases performance by two orders of magnitude on satisfiable instances,
    whilst not affecting performance on unsatisfiable instances. This algorithm can also be
    parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
    instances, we get a further speedup of over thirty from thirty-six cores, or ninety five from
    ten distributed-memory hosts. Finally, we show that solution-biased search is also suitable for
    optimisation problems, by using it to improve two maximum common induced subgraph algorithms.
\end{abstract}

@@

This new form of search can also be parallelised, with a \emph{much} simpler implementation than
conventional work-stealing. By running many threads with different random seeds but the same restart
schedule, and sharing nogoods only following restarts, we can achieve speedups of thirty-one from a
thirty-six core machine, or ninety-five by using ten such machines with message passing.

\subsection{Background}

@@

@@

\Cref{figure:parallel} shows how this scheme performs in practice. Sequentially, we can solve 14,357
instances within the thousand second timeout, with the last instance being solved at 939.0 seconds.
Using thirty-six threads on machines with two eighteen core processors, we can solve 14,357
instances with a timeout of only 74.2 seconds, giving an aggregate speedup
\cite{DBLP:conf/cpaior/HoffmannMNPRS018} of 12.7.

@@

Because we are using nogood recording, an alternative approach is possible. Rather than using the
Luby sequence for restarts, we could restart after either a constant number of backtracks, or simply
after a certain time interval has passed---this bounds the maximum possible idle time that threads
could spend at a barrier. \Cref{figure:parallel} also shows the effects of restarting every 100ms.
Sequentially, this approach is slightly better than using the Luby sequence, being able to solve
14,370 instances, with the last at 996.4 seconds. With thirty-six threads, solving this many
instances takes 31.8 seconds, giving an aggregate speedup of 31.4.
